\subsection{Radiation transport: Flux-limited diffusion}
\label{sec.num.rad-fld}

In addition to the ray-tracing approach for radiation transport
described in Section~\ref{sec.num.raytracing}, \enzo\ currently
includes a field-based radiation transport solver for problems posed
on uniform (i.e.~non-AMR, non-static mesh refinement) grids, which has
been tuned for large-scale simulations involving many ionizing
sources.  Detailed explanations of the model and solution approach may
be found in \citet{NBHBROW2007},
\citet{ReynoldsHayesPaschosNorman2009}, \citet{NRS2009}, and
\citet{2013arXiv1306.0645N}, the salient features of which are
reproduced here.  In addition, comparisons of this solver with other
astrophysical radiation transport solvers may be found in
\cite{IlievEtAl2009}.  \enzo's field-based radiation solver focuses on
a flux-limited diffusion approximation for cosmological radiative
transfer, with couplings to both the gas energy and chemical number
densities.


The system of equations (\ref{eq:fld_radiation}-\ref{eq:fld_heating})
along with the chemical network (Equation~\ref{eq:species_evolution})
is solved independently of \enzo's hydrodynamics, gravity and
dark-matter solvers (Sections \ref{sec.hydro.ppm}-\ref{sec.ov.nbody}),
thereby allowing the advective portions of Equations
(\ref{eq:fld_radiation}) and (\ref{eq:species_evolution}) to be taken
care of by the fluid solvers.  Due to the disparate time scales
between radiation transport and chemical ionization and heating, the
remainder of these equations is solved using an operator-split
algorithm.  Within a given timestep to evolve $(E_r^n, e_c^n, {\mathrm
n}_i^n) \to (E_r^{n+1}, e_c^{n+1}, {\mathrm n}_i^{n+1})$, we first
evolve equation (\ref{eq:fld_radiation}): $(E_r^n, e_c^n, {\mathrm
n}_i^n) \to (E_r^{n+1}, e_c^{n}, {\mathrm n}_i^{n})$.  This uses an
implicit Euler time discretization, and a second-order centered finite
difference spatial discretization, resulting in a large linear system
of equations.  These are solved using a preconditioned conjugate
gradient iteration, where the preconditioner consists of a geometric
multigrid solver.  Both of these linear solvers are provided by the
HYPRE linear solver library \citep[see][]{FalgoutYang2002,
hypre-manual}.

We then evolve the heating and chemistry system,
Equations~(\ref{eq:fld_heating}) and (\ref{eq:species_evolution}):
$(E_r^{n+1}, e_c^n, {\mathrm n}_i^n) \to (E_r^{n+1}, e_c^{n+1},
{\mathrm n}_i^{n+1})$.  Due to the lack of spatial derivatives (since
advection is handled elsewhere), this system is a coupled system of
nonlinear ordinary differential equations.  This utilizes an implicit
quasi-steady-state approximation, formulated as follows.  We consider
the modified equations,
\begin{eqnarray}
  \label{eq:fld_heating_qss}
  \frac{\partial e_c}{\partial t} &=& -\frac{2\dot{a}}{a} e_c +
    \Gamma\left(\bar{E}_r,\bar{\mathrm n}_i\right) - 
    \Lambda\left(\bar{E}_r,\bar{\mathrm n}_i\right), \\
  \label{eq:fld_chemistry_qss}
  \frac{\partial {\mathrm n}_i}{\partial t} &=& k_{i,j}\left(\bar{e}_c\right)
    {\mathrm n}_e \bar{\mathrm n}_j - {\mathrm n}_i 
    \Gamma_i^{ph}\left(\bar{E}_r\right), \quad i=1,\ldots,N_s,
\end{eqnarray}
where we have defined the time-centered ``background'' states
$\bar{E}_r = \left(E_r^{n}+E_r^{n+1}\right)/2$, 
$\bar{\mathrm n}_i = \left({\mathrm n}_i^{n}+{\mathrm n}_i^{n+1}\right)/2$
and $\bar{e}_c = \left(e_c^{n}+e_c^{n+1}\right)/2$.  These equations
may each be solved analytically for their solution at the time
$t$, which we denote by
\begin{eqnarray}
  \label{eq:fld_heating_qss_sol}
  e_c(t) &=& \text{sol}_{e}\left(t,\bar{E}_r,\bar{\mathrm n}_i,e_c^n\right), \\
  \label{eq:fld_chemistry_qss_sol}
  {\mathrm n}_i(t) &=& \text{sol}_{\mathrm n_i}
  \left(t,\bar{E}_r,\bar{e_c},\mathrm n_i^n\right), \quad i=1,\ldots,N_s. 
\end{eqnarray}
We then define a nonlinear system of equations to compute the
time-evolved solutions $\left(E_r^{n+1}, e_c^{n+1}, 
{\mathrm n}_i^{n+1}\right)$ as
\begin{eqnarray}
  \label{eq:fld_heating_qss_fe}
  f_e(e_c^{n+1},{\mathrm n}_i^{n+1}) &\equiv& e_c^{n+1} -
    \text{sol}_{e}\left(t^{n+1},\bar{E}_r,\bar{\mathrm n}_i,e_c^n\right)
    = 0, \\
  \label{eq:fld_chemistry_qss_fn}
  f_{\mathrm n_i}(e_c^{n+1},{\mathrm n}_i^{n+1}) &\equiv& 
    {\mathrm n}_i(t) - \text{sol}_{\mathrm n_i}
    \left(t,\bar{E}_r,\bar{e_c},\mathrm n_i^n\right)=0, \quad
    i=1,\ldots,N_s.  
\end{eqnarray}
This system of $N_s+1$ nonlinear equations is solved using a damped
fixed point iteration, 
\[
   U_i = U_i - \lambda f_i(U), \quad i=1,\ldots,N_s+1,
\]
where $U$ is a vector containing the solutions to equations
(\ref{eq:fld_heating_qss_fe}-\ref{eq:fld_chemistry_qss_fn}).  In this
iteration, for the first 50 sweeps we use $\lambda=1$. For more
challenging problems where this does not converge, we switch to a
damping parameter of $\lambda=0.1$.
