\subsubsection{The \enzo\ test suite}
\label{sec.tests.suite}

\enzo\ is capable of simulating a large variety of problems types,
with all but a few of these types requiring only a parameter file as
an input.  The most notable exception is the cosmology simulation,
which takes as input initial conditions created by other codes.  The
suite of test problems spans a wide range in complexity.  At one end
of this spectrum are simple problems that utilize only a single
component of \enzo\ and for which analytic solutions exist for
comparison with the simulation results.  At the opposite end are
problems that exercise a large portion of \enzo's machinery in
concert.  Together, the available problem types fully cover all of
\enzo's functionality.  This enables them to serve as a vehicle for
verifying that the code's behavior remains stable over time, on new
computing platforms, and after modification of the codebase.

\enzo\ uses an automated testing framework that allows a user to run,
with a single command, a set of test problems and compare the results
against results produced by any other version of the code.  Within the
\enzo\ source distribution, the test problem parameter files are
stored in a nested directory structure organized according to the
primary functionality tested (e.g., hydrodynamics, gravity, cooling,
etc.).  Each parameter file is accompanied by a text file containing
various descriptive keywords, such as the machinery tested, the
dimensionality, and the approximate run time.  A test runner script is
responsible for taking as input from the user a series of keywords
that are used to select a subset of all available test problems.  The
test problems are also grouped into three suites: the quick, push, and
full suites, each a superset of the ones named before.  The quick
suite is considered to minimally cover the primary functionality in
\enzo\ and is designed to be run repeatedly during the development
process.  The push suite has slightly increased feature coverage and
is mandated to be run before code changes are accepted into the main
repository.  The full suite consists of all test problems that can be
run with no additional input.  Approximate run times for the quick,
push, and full suite are 15 minutes, 1 hour, and 60 hours,
respectively, on a relatively new desktop computer (circa 2013).

After the test problems are selected by the test runner script, they
are run in succession using either the \enzo\ executable contained
within that distribution or an external executable built from another
\enzo\ version and specified by the user.  This allows for any version
of the code to be tested with an identical set of test problems.
After running the test problem simulations, the test runner then
performs a series of basic analysis tasks using the \texttt{yt}
analysis toolkit \citep{2011ApJS..192....9T, 2011arXiv1112.4482T}.
The default analysis performed on all test problems includes
calculation of various statistics (such as extrema, mean, and
variance) on the fields present in the output data.  Custom analysis
provided by scripts that accompany the test problem parameter files is
run for special cases, such as when an analytical solution exists that
can be compared against the simulation data.  After the analysis is
performed, the results are compared against a set of gold standard
results that are maintained on a website and downloaded on the fly by
the test runner script.  Alternately, results from any version of
\enzo\ can be stored locally and compared to any other version of the
code.  In Section~\ref{sec.tests.problems}, we describe some of the
key test problems that are used to verify proper behavior.  All of
these test problems, as well as the scripts to generate the figures
from them, are available at
\url{http://bitbucket.org/enzo/enzo-method-paper}, the Bitbucket
repository for the \enzo\ method paper.  The \texttt{test\_problems}
subdirectory in this repository contains all of the files necessary to
regenerate all of the figures in Section~\ref{sec.tests.problems}.  We
note for completeness that the figures in
Section~\ref{sec.tests.problems} were generated using \texttt{yt}
version 2.5.3 and the \url{https://bitbucket.org/enzo/enzo-dev}
repository of \enzo\ with changeset \texttt{bad70b86fcd6}, which
corresponds to the \enzo\ 2.3 code release.

\subsubsection{Code comparisons}
\label{sec.tests.compare}

Over the course of its existence, \enzo\ has been involved in numerous
comparisons with other astrophysical codes used for self-gravitating
fluid dynamics.  In general, \enzo\ behaves in a manner similar to
other grid-based (and particularly AMR-based) codes, as we will
summarize below.

\enzo\ has been involved in multiple cosmological code comparisons,
including the Santa Barbara Cluster code comparison project
\citep{SantaBarbara}, a large comparison of N-body simulations
\citep{2008CS&D....1a5003H}, as well as several direct comparisons
between \enzo\ and the GADGET SPH code in a variety of set-ups,
including N-body and adiabatic hydrodynamics
\citep{2005ApJS..160....1O,2005MNRAS.364..909V, 2011MNRAS.418..960V}
and simulations of the Lyman-alpha forest \citep{2007MNRAS.374..196R}.
Compared to the other codes involved in these projects, the \enzo\
code typically has a more difficult time resolving small-scale
self-gravitating structures (for an equivalent dark matter particle
count and nominally equivalent force resolution), but does comparably
well as a tree-based code for larger structure, and is typically
superior in terms of resolving fluid features due to its higher-order
(and artificial viscosity-free) PPM hydrodynamics solver.  When
examining classical test cases such as the Santa Barbara project
\citep{SantaBarbara}, \enzo\ forms galaxy clusters with very similar
density, temperature, and entropy profiles to other grid-based codes
that use Godunov-type hydro methods, which systematically differed
from particle-based codes using SPH in this code comparison.
Similarly, in galaxy cluster simulations that look specifically at the
properties of cosmological shocks \citep[e.g.][]{2011MNRAS.418..960V},
\enzo\ produces results that are similar to other high-order
grid-based hydrodynamics codes, and a far superior performance of
\enzo\ is observed (in terms of resolution of fluid features and shock
detection) in low-density regions when compared to a particle-based
code.  In tests of the Lyman-alpha forest that include radiative
cooling and a uniform metagalactic ultraviolet background, \enzo\ and
GADGET provide results on metrics such as the matter power spectrum
that are comparable to within 5\% \citep{2007MNRAS.374..196R}.

Several comparisons have been made that focus specifically on
hydrodynamics solvers and fluid behavior.  In particular, the work of
\citet{2007MNRAS.380..963A} and \citet{Tasker2008} perform direct
comparisons between several grid- and particle-based codes for a
variety of fluid-centric test problems (including shocked gas clouds,
self-gravitating, translating clouds, and Sedov-Taylor explosions),
and show that \enzo\ is comparable or superior in behavior to the
other grid-based hydrodynamics codes involved in the comparison, and
provide useful information on the sort of practical challenges that a
user of an AMR code such as \enzo\ may experience.  More specific
comparisons, including one testing the linear and nonlinear growth of
the Kelvin-Helmholz instability \citep{2012ApJS..201...18M}, as well
as a comparison that more broadly examines Galilean invariance in
grid-based codes \citep{2010MNRAS.401.2463R}, show that \enzo, and in
particular its implementation of the PPM hydrodynamics solver,
converge to the correct solution as expected, and generally provide
less diffusive solutions than lower-order codes, including those that
use artificial viscosity.  Finally, there have been two code
comparison projects that focus on turbulence simulations.  The first
studied the behavior of decaying isothermal supersonic turbulence
\citep{2009A&A...508..541K}, and the second examined supersonic
magnetohydrodynamical turbulence \citep{2011ApJ...737...13K}.  Both
included the \enzo\ code, with the former testing the PPM
hydrodynamics and the latter both the constrained transport MHD
implementation of \citet{Collins10} and the Dedner method of
\citet{WangAbelZhang08}.  In both cases, \enzo\ performed similarly to
other grid-based codes that use Godunov-based fluid solvers, and
typically had better effective resolution than particle-based codes
when using the same number of particles as the number of grid cells in
the \enzo\ simulation.

Three other comparisons between the \enzo\ code and other simulation
tools have been performed that focus on physics other than gravity and
fluid flow.  The flux-limited diffusion radiation transport scheme was
measured against several test problems by \citet{IlievEtAl2009}, which
involved tests with and without analytical solutions.  \enzo\ produced
results similar to both the analytical solutions and results obtained
by other codes.  We note, however, that there were minor differences
throughout the comparison between various codes, and the majority of
the codes differed in at least a subset of the tests.
\citet{2011ApJ...726...55T} show the result of varying reaction rate
coefficients for the formation of molecular hydrogen via the 3-body
process in both the \enzo\ and GADGET-2 codes, observing similar
trends between the two codes. However, at nominally equivalent
resolution where the particle and cell gas masses are comparable,
\enzo\ simulations typically displayed a substantially higher level of
gas structure.  This is unsurprising due to \enzo's higher-order hydro
solver.  Finally, \citet{2012ApJ...744...52P} show the results of
comparing \enzo\ in its non-AMR mode to a smoothed-particle
hydrodynamics code, SNSPH, in the context of common-envelope binary
stellar interactions.  The authors show that the codes display
reasonable convergence properties as a function of simulation
resolution, and also agree quite well with each other. However, the
observed mass-loss rates do not agree particularly well with
observations.
